---
title: "Exercise_11"
output: html_document
date: "2024-04-01"
---
Before we get started, we will load all the packages needed and load the data set.
```{r}
new_pkg <- c("tidyverse", "dplyr", "mosaic", "ggplot2", "infer", "skimr", "cowplot", "broom")
not_installed <- new_pkg[!new_pkg %in% rownames(installed.packages())]
if (length(not_installed) > 0) install.packages(not_installed) 

library("tidyverse")
library("dplyr")
library("mosaic")
library("ggplot2")
library("infer")
library("skimr")
library("cowplot")
library("broom")
```

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/AVONETdataset1.csv"
d <- read_csv(f) %>%
  select(Species1, Family1, Order1, Beak.Length_Culmen, Beak.Width, Beak.Depth, Tarsus.Length, Wing.Length, Tail.Length, Mass, Habitat, Migration, Trophic.Level, Trophic.Niche, Min.Latitude, Max.Latitude, Centroid.Latitude, Primary.Lifestyle, Range.Size)

skim(d)
```
The output indicates 7 variables are categorical and 12 are numeric. They were listed in the following output. 
  character                7     
  numeric                  12   

Challenge 1
Step 1
```{r}
d_1 <- d %>%
  mutate(Migration = as.factor(Migration)) %>%
  mutate(logMass = log(Mass))

skim(d_1)
```
We use skim() to confirm that Migration variable is successfully converted to a factor variable.

```{r}
p1 <- ggplot(drop_na(d_1, Trophic.Level), aes(Trophic.Level, logMass)) +
    geom_boxplot()
p2 <- ggplot(drop_na(d_1, Migration), aes(Migration, logMass)) +
    geom_boxplot()

plot_grid(p1,p2)
```

Step 2
```{r}
m_TL <- lm(logMass ~ Trophic.Level, data = d_1)
summary(m_TL)
```

```{r}
m_M <- lm(logMass ~ Migration, data = d_1)
summary(m_M)
```
In both summaries, in the global test of significance, the F statistics are large enough to reject the null hypothesis of an F value of zero. Variable log(Mass) is associated with both categories. Both Migration 2 and Migration 3 are different from the reference level (Migration 1)

```{r}
d_1$Migration <- relevel(d_1$Migration, ref = "2")
m_M_relevel <- lm(logMass ~ Migration, data = d_1)
summary(m_M_relevel)
d_1$Migration <- relevel(d_1$Migration, ref = "1")
```
The missing pair Migration2 and Migration3 are also different. The last line in this chunk is for dataset use in other chunks (e.g., in previous model running).

Step 3
```{r}
m_M_aov <- aov(logMass ~ Migration, data = d_1)
posthoc <- TukeyHSD(m_M_aov, which = "Migration", ordered = TRUE, conf.level = 0.95)
posthoc
```
All categories are significantly different from one another.

Step 4
```{r}
m_TL_aov <- aov(logMass ~ Trophic.Level, data = d_1)
```


```{r}
original.F <- m_TL_aov %>%
  tidy() %>%
  filter(term=="Trophic.Level")
original.F
```

```{r}
permuted.F <- d_1 %>%
  specify(logMass ~ Trophic.Level) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "F")

visualize(permuted.F) +
  shade_p_value(obs_stat=original.F$statistic, direction="greater")

```

```{r}
(p.value <- permuted.F %>%
  get_p_value(obs_stat = original.F$statistic, direction="greater"))
```

Challenge 2
Step 1
In the following chunks, two coefficients are calculated and named by themselves.
```{r}
d_2 <- d_1 %>%
  mutate(logBL = log(Beak.Length_Culmen)) %>%
  mutate(logTaL = log(Tarsus.Length))

```

```{r}
m_BL <- lm(logBL ~ logMass, data = d_2)
m_TaL <- lm(logTaL ~ logMass, data = d_2)
```

```{r}
d_2 <- d_2 %>%
  mutate(relBL = m_BL$residuals) %>%
  mutate(relTaL = m_TaL$residuals)
```
For convenience's sake, I didn't directly use the name provided in the prompt word-by-word (Relative beak length, Relative tarsus length)

Step 2
```{r}
p3 <- ggplot(d_2, aes(Primary.Lifestyle, relTaL)) +
    geom_boxplot()
p4 <- ggplot(d_2, aes(Trophic.Niche, relBL)) +
    geom_boxplot()

plot_grid(p3,p4)
```

Step 3
```{r}
d_3 <- d_2 %>%
  drop_na(Migration)

hist(d_3$Range.Size)
```
The shape indicates that we had better log transform the variable.

```{r}
d_3 <- d_3 %>%
  mutate(logRS = log(Range.Size))
```

```{r}
m_RSM_aov <- aov(logRS ~ Migration, data = d_3)
summary(m_RSM_aov)
```
Based on the global model, logRS is associated with form of migration. We also need to calculate the “R-squared value.”

```{r}
SSR <- sum((m_RSM_aov$fitted.values - mean(m_RSM_aov$model$logRS))^2)
SSR
```

```{r}
SSY <- sum((m_RSM_aov$model$logRS - mean(m_RSM_aov$model$logRS))^2)
SSY
```

```{r}
rsq <- SSR/SSY
rsq
```

About 8.69% of the variance in the measure of range size is associated with Migration behavior style.

```{r}
m_RSM <- lm(logRS ~ Migration, data = d_3)
summary(m_RSM)
```
The reference level is Migration1. Both Migration2 and Migration3 are different from Migration1. The Multiple R-squared value in the lm model is the same with R-squared value calculated from aov model.

```{r}
d_3$Migration <- relevel(d_3$Migration, ref = "2")
m_RSM_relevel <- lm(logRS ~ Migration, data = d_3)
summary(m_RSM_relevel)
d_3$Migration <- relevel(d_3$Migration, ref = "1")
```
The missing pair Migration2 and Migration3 are also different. Again we just relevel the variable back.

```{r}
posthoc_3 <- TukeyHSD(m_RSM_aov, which = "Migration", ordered = TRUE, conf.level = 0.95)
posthoc_3

```
All categories are significantly different from one another.

Step 4
We will directly filter from d_2 as the dataset where relBL is already available without any unwanted drop_na().
```{r}
d_4 <- d_2 %>%
  filter(Order1 == "Passeriformes") %>%
  mutate(PLTL = interaction(Primary.Lifestyle, Trophic.Level))
```

Here are some boxplots.
```{r}
p5 <- ggplot(d_4, aes(Primary.Lifestyle, relBL)) +
    geom_boxplot()
p6 <- ggplot(d_4, aes(Trophic.Level, relBL)) +
    geom_boxplot()
p7 <- ggplot(d_4, aes(PLTL, relBL)) +
    geom_boxplot() +
  theme(axis.text.x = element_text(size = 5,angle = 270))

plot_grid(p5,p6,p7)
```
```{r}
m_relBLPL <- lm(relBL ~ Primary.Lifestyle, data = d_4)
summary(m_relBLPL)
```


```{r}
m_relBLPL_aov <- aov(relBL ~ Primary.Lifestyle, data = d_4)
summary(m_relBLPL_aov)
```

```{r}
m_relBLTL <- lm(relBL ~ Trophic.Level, data = d_4)
summary(m_relBLTL)
```

```{r}
m_relBLTL_aov <- aov(relBL ~ Trophic.Level, data = d_4)
summary(m_relBLTL_aov)
```
relBL is associated with both PL and TL. In each aov() model, PL or TL yields a main effect. PL and TL explain about 6% and 5% of total variance respectively in their separate one-way models.

Step 5
```{r}
m_5 <- aov(relBL ~ Primary.Lifestyle + Trophic.Level, data = d_4)
summary(m_5)
```
There is still a main effect of PL when taking TL into account and there is a main effect of TL when we take PL into account.

Step 6
```{r}
m_5 <- aov(relBL ~ Primary.Lifestyle + Trophic.Level + Primary.Lifestyle:Trophic.Level, data = d_4)
summary(m_5)
```
It looks like there is indeed a significant main effect of each variable (TL; PL) as well as an interaction between the two.

Step 7
```{r}
interaction.plot(
  x.factor = d_4$Primary.Lifestyle,
  xlab = "Primary.Lifestyle",
  trace.factor = d_4$Trophic.Level,
  trace.label = "Trophic.Level",
  response = d_4$relBL,
  fun = base::mean,
  ylab = "Mean Relative beak length"
)
```

Additional Steps
There were multiple models in this assignment. We will take Challenge 2 Step 4 to Step 7 model groups as examples (on "Passeriformes"). As noted in Module 20, a general rule of thumb for “equal” variances is to compare the largest and smallest within-grouping level standard deviations and, if this value is less than 2.
```{r}
stats <- d_4 %>% 
  group_by(Primary.Lifestyle, Trophic.Level) %>%
  summarize("mean(relBL)" = mean(relBL),
    "sd(relBL)"= sd(relBL))
(max(stats$`sd(relBL)`)/min(stats$`sd(relBL)`))
```

If we directly jump to two-variable ANOVA and keep the rule of thumb "less than 2," the dataset doesn't pass the check. However, if we check two variables separately the stories are different.

```{r}
stats_PL <- d_4 %>% 
  group_by(Primary.Lifestyle) %>%
  summarize("mean(relBL)" = mean(relBL),
    "sd(relBL)"= sd(relBL))
(max(stats_PL$`sd(relBL)`)/min(stats_PL$`sd(relBL)`))
```

```{r}
stats_TL <- d_4 %>% 
  group_by(Trophic.Level) %>%
  summarize("mean(relBL)" = mean(relBL),
    "sd(relBL)"= sd(relBL))
(max(stats_TL$`sd(relBL)`)/min(stats_TL$`sd(relBL)`))
```

Here are some other visual checks for PL:
```{r}
p8 <- ggplot(d_4, aes(relBL))+
  geom_histogram()+
  facet_wrap(~Primary.Lifestyle)
p9 <- ggplot(d_4, aes(relBL))+
  geom_histogram()+
  facet_wrap(~Trophic.Level)
p10 <- ggplot(d_4, aes(relBL))+
  geom_histogram()+
  facet_wrap(~PLTL)

plot_grid(p8,p9,p10)
```

```{r}
p11 <- ggplot(m_relBLPL_aov, aes(m_relBLPL_aov$residuals))+
  geom_histogram()+
  facet_wrap(~Primary.Lifestyle)
p12 <- ggplot(m_relBLTL_aov, aes(m_relBLTL_aov$residuals))+
  geom_histogram()+
  facet_wrap(~Trophic.Level)


plot_grid(p11,p12)
```

```{r}
p13 <- ggplot(m_5, aes(m_5$residuals))+
  geom_histogram()+
  facet_wrap(~Primary.Lifestyle+Trophic.Level)

p13
```











