---
title: "Exercise_11"
output: html_document
date: "2024-04-01"
---
Before we get started, we will load all the packages needed and load the data set.
```{r}
new_pkg <- c("tidyverse", "dplyr", "mosaic", "ggplot2", "infer", "skimr", "cowplot", "broom")
not_installed <- new_pkg[!new_pkg %in% rownames(installed.packages())]
if (length(not_installed) > 0) install.packages(not_installed) 

library("tidyverse")
library("dplyr")
library("mosaic")
library("ggplot2")
library("infer")
library("skimr")
library("cowplot")
library("broom")
```

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/AVONETdataset1.csv"
d <- read_csv(f) %>%
  select(Species1, Family1, Order1, Beak.Length_Culmen, Beak.Width, Beak.Depth, Tarsus.Length, Wing.Length, Tail.Length, Mass, Habitat, Migration, Trophic.Level, Trophic.Niche, Min.Latitude, Max.Latitude, Centroid.Latitude, Primary.Lifestyle, Range.Size)

skim(d)
```
The output indicates 7 variables are categorical and 12 are numeric. They were listed in the following output. 
  character                7     
  numeric                  12   

Challenge 1
Step 1
```{r}
d_1 <- d %>%
  mutate(Migration = as.factor(Migration)) %>%
  mutate(logMass = log(Mass)) %>%
  drop_na(Trophic.Level, Migration)

skim(d_1)
```
We use skim() to confirm that Migration variable is successfully converted to a factor variable.

```{r}
p1 <- ggplot(d_1, aes(Trophic.Level, logMass)) +
    geom_boxplot()
p2 <- ggplot(d_1, aes(Migration, logMass)) +
    geom_boxplot()

plot_grid(p1,p2)
```

Step 2
```{r}
m_TL <- lm(logMass ~ Trophic.Level, data = d_1)
summary(m_TL)
```

```{r}
m_M <- lm(logMass ~ Migration, data = d_1)
summary(m_M)
```
In both summaries, in the global test of significance, the F statistics are large enough to reject the null hypothesis of an F value of zero. Variable log(Mass) is associated with both categories. Both Migration 2 and Migration 3 are different from the reference level (Migration 1)

```{r}
d_1$Migration <- relevel(d_1$Migration, ref = "2")
m_M_relevel <- lm(logMass ~ Migration, data = d_1)
summary(m_M_relevel)
```
The missing pair Migration2 and Migration3 are also different.

Step 3
```{r}
m_M_aov <- aov(logMass ~ Migration, data = d_1)
posthoc <- TukeyHSD(m_M_aov, which = "Migration", ordered = TRUE, conf.level = 0.95)
posthoc
```
All categories are significantly different from one another.

Step 4
```{r}
m_TL_aov <- aov(logMass ~ Trophic.Level, data = d_1)
```


```{r}
original.F <- m_TL_aov %>%
  tidy() %>%
  filter(term=="Trophic.Level")
original.F
```

```{r}
permuted.F <- d_1 %>%
  specify(logMass ~ Trophic.Level) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "F")

visualize(permuted.F) +
  shade_p_value(obs_stat=original.F$statistic, direction="greater")

```

```{r}
(p.value <- permuted.F %>%
  get_p_value(obs_stat = original.F$statistic, direction="greater"))
```

Challenge 2
Step 1
In the following chunks, two coefficients are calculated and named by themselves.
```{r}
d_2 <- d_1 %>%
  mutate(logBL = log(Beak.Length_Culmen)) %>%
  mutate(logTaL = log(Tarsus.Length)) %>%

```

```{r}
m_BL <- lm(logBL ~ logMass, data = d_2)
m_TaL <- lm(logTaL ~ logMass, data = d_2)
```

```{r}
d_2 <- d_2 %>%
  mutate(relBL = m_BL$residuals) %>%
  mutate(relTaL = m_TaL$residuals)
```
For convenience's sake, I didn't directly use the name provided in the prompt word-by-word (Relative beak length, Relative tarsus length)

Step 2
```{r}
p3 <- ggplot(d_2, aes(Primary.Lifestyle, relTaL)) +
    geom_boxplot()
p4 <- ggplot(d_2, aes(Trophic.Niche, relBL)) +
    geom_boxplot()

plot_grid(p3,p4)
```

Step 3
```{r}
betas_ct <- lm_radiation("Catarrhini")
```

```{r}
betas_pl <- lm_radiation("Platyrrhini")
```


```{r}
betas_st <- lm_radiation("Strepsirhini")
```

Beta0s (intercepts) are highly different; those of platyrrhines and strepsirhines share relatively closer intercepts. Platyrrhines and strepsirhines share relatively similar beta1s (slopes), whereas they are different from that of catarrhines.

Step 6
```{r}
d6 <- d3 %>%
  mutate(ECVhat = beta0 + beta1 * Group_size) %>%
  mutate(residuals = ECV -ECVhat)
str(d6)
```

SE for the slope coefficient by hand:
```{r}
(df_regression <- 1) 
(df_error <- nrow(d6) - df_regression - 1) 
```

```{r}
SSE <- sum((d6$residuals)^2)
SSE
```

```{r}
(MSE <- SSE/df_error) 
```

```{r}
SSX <- sum((d6$Group_size - mean(d6$Group_size))^2)
SEbeta1 <- sqrt(MSE/SSX)
SEbeta1

```

SE for the slope coefficient extracted from the model:
```{r}
m.summary <- tidy(m)
tidy(m)$std.error[2]
tidy(m)$p.value[2]
```
In the chunk above, the p value is extracted as well.

The 95% CI by hand:
```{r}
alpha <- 0.05
```

```{r}
(lower <- beta1 - qt(1 - alpha/2, df = nrow(d6) - 2) * SEbeta1)
(upper <- beta1 + qt(1 - alpha/2, df = nrow(d6) - 2) * SEbeta1)
```

The 95% CI extracted from model:
```{r}
CI <- confint(m, level = 1 - alpha)
CI[2,1:2]
```

p value calculated by hand:
```{r}
calc.statistic <- (beta1 - 0)/SEbeta1
(p <- 2 * pt(calc.statistic, df = nrow(d6) - 2, lower.tail = FALSE))
```
p value was already extracted earlier.

Step 8
We need to permute slope coefficient estimates. The sentence from Module 18 explain what we need to permute: "essentially, each permutation involves randomly shuffling values of the response variable under the null hypothesis that they are independent of the explanatory variable."
```{r}
alpha <- 0.05
confidence_level <- 1 - alpha
p_lower <- alpha/2
p_upper <- 1 - (alpha/2)
degrees_of_freedom <- nrow(d6) - 2
critical_value <- qt(p_upper, df = degrees_of_freedom)

original.slope <- m %>%
tidy(conf.int = TRUE, conf.level = confidence_level) %>%
mutate(lower = estimate - std.error * critical_value, upper = estimate + std.error *
    critical_value) %>%
    filter(term == "Group_size")
original.slope 
```

```{r}
permuted.slope <- d6 %>%
  specify(ECV ~ Group_size) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "slope")

head(permuted.slope)
```

```{r}
permuted.slope.summary <- permuted.slope %>%
  summarize(
    estimate = mean(stat)
  )

permuted.slope.summary
```

```{r}
(p.value <- permuted.slope %>%
  get_p_value(obs_stat = original.slope$estimate, direction="two_sided"))
```
The p value is 0.

Step 8
```{r}
boot.slope <- d6 %>%
  specify(ECV ~ Group_size) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "slope")

head(boot.slope)
```
```{r}
boot.slope.summary <- boot.slope %>%
  summarize(
    estimate = mean(stat)
  )

boot.slope.summary
```

```{r}
(CI.percentile <- get_ci(boot.slope, level = 1 - alpha, type = "percentile"))
(CI.theory <- get_ci(boot.slope, level = 1 - alpha, type = "se", point_estimate = pull(boot.slope.summary,
    estimate)))
```

These CIs suggest that beta1 (slope coefficient) is different from zero.
